---
title: "Joining Review Exercises"
author: "Your Name Here"
date: "`r format(Sys.time(), '%B %d, %Y')`"
format:
  html:
    self-contained: true
    toc: true
    toc-location: left
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
# asfdsadf
```


# Load the datasets

```{r}

counties_demographics <- readRDS(here("data", "counties_demographics.rds"))
counties_elexresults <- readRDS(here("data", "counties_elexresults.rds"))
counties_rural_urban_suburban <- readRDS(here("data", "counties_rural_urban_suburban.rds"))



```

# Exercises

Join together the election results and urban/suburban/rural tables above, and then use the merged table to:

-   filter for just Virginia

-   group/summarize to calculate the mean pct of Trump 2020 vote by urban/rural categories

```{r}
merged_data <- inner_join(counties_elexresults, counties_rural_urban_suburban, by = "geoid")

merged_data_va <- merged_data %>% 
  filter(state_code == "VA")

merged_data_va_avg_pct_r <- merged_data_va %>%
  group_by(rural_urban) %>% 
  summarise(avg_pct_r = mean(pct_r_2020))
  
merged_data_va_avg_pct_r

```

Take the grouped dataset, and create a column chart using ggplot to display it

```{r}

ggplot(merged_data_va_avg_pct_r, aes(x = reorder(rural_urban, avg_pct_r), y = avg_pct_r)) + 
  geom_col() + 
  scale_y_continuous(name = "Avg pct of Trump 2020 Vote", labels = scales::comma) +
  scale_x_discrete(name = "Urban/Rural Category") 

```

Now using the joined dataset, come up with two groupings you find interesting or newsworthy in a particular state, and replicate the steps above them them instead.

(i.e. instead of for urban/rural, perhaps you want to look at pct of white college grads in Michigan, etc. If you're having trouble, you can stick with Virginia as your state.)

```{r}

merged_data <- inner_join(counties_elexresults, counties_rural_urban_suburban, by = "geoid")

merged_data_il <- merged_data %>% 
  filter(state_code == "IL")

merged_data_il_margin_2020 <- merged_data_il %>%
  group_by(rural_urban) %>% 
  summarise(avg_margin_2020 = mean(margin_2020))
  
merged_data_il_margin_2020


```


```{r}

ggplot(merged_data_il_margin_2020, aes(x = reorder(rural_urban, avg_margin_2020), y = avg_margin_2020)) + 
  geom_col() + 
  scale_y_continuous(name = "Avg Margin in 2020", labels = scales::comma) +
  scale_x_discrete(name = "Urban/Rural Category") 

```
